---
title: "Predicting Trimester Using Cephalic Index"
author: "Usman"
date: "2023-12-06"
output: 
  html_document: 
    toc: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# *R Markdown of Cephalic Index Modelling for Predicting Human Trimester*

## **Inspecting data for Multicolinearity using 'corrplot' package**


```{r Cephalic Index}
library(dplyr)
library(readxl)
cephalic <- read_excel("~/Bluetooth Exchange Folder/sagagi2.xlsx",sheet = "Sheet2")
cephalic

# logistic regression analysis
cephalic$sex<-cephalic$SEX
cephalic$logit<-cephalic$TRIMESTER
cephalic[cephalic$logit=="2ND"|cephalic$logit=="2nd"|cephalic$logit=="2NDF",]$logit<-"0"
cephalic[cephalic$logit=="3RD"|cephalic$logit=="3rd",]$logit<-"1"
cephalic[cephalic$sex=="F",]$sex<-"0"
cephalic[cephalic$sex=="M",]$sex<-"1"
#glm(logit~sex, data=cephalic,family = "binomial")
cephalic$logit<-as.factor(cephalic$logit)

library(corrplot)
library(Hmisc)
cor <- rcorr(as.matrix(cephalic[,-c(1,2,8:11)]))
# To produce the correlation coefficience
cor$r %>% as.table()

# or alternatively use "psych"
library(psych)
# To provide with a matrix table
pairs.panels(cephalic[,-c(1,2,8:11)])
```

## **Splitting data by into Training (60%) and Testing (40%)**
 
 ### **Using the base R 'sample_frac' function or 'createDataPartition' of \n "caret" packages**
 
```{r}
library(caret)
set.seed(123)
training.samples <- cephalic$logit %>% 
  createDataPartition(p = 0.6, list = FALSE)
train.data  <- cephalic[training.samples, ]
test.data <- cephalic[-training.samples, ]
nrow(train.data)
nrow(test.data)

# Using the "sample_frac" to divide our data into Training and Testing data set
train.data1 <- sample_frac(cephalic, 0.6)
test.data1 <- sample_frac(cephalic, 0.4)
nrow(train.data1)
nrow(test.data1)
```
 
## Modelling Using Cephalic Index and Biparietal Diameter (BPD)

```{r}
model <- glm( logit~BPD+CI, 
              data = train.data1, family = binomial)

probabilities <- model %>% predict(test.data1, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "3rd", "2nd")
```


## Logistic Regression S curve plot

```{r}
# A Logistic Regression Line Curve for Testing and Training Data Set


A<-test.data1 %>%
  mutate(prob = ifelse(logit == "1", 1, 0)) %>%
  ggplot(aes(CI, prob)) +geom_jitter(height = 0.01)+
  geom_point(alpha = 0.2) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  labs(
    title = "Logistic Regression Model \n Testing data set 40%", 
    x = "Cephalic Index",
    y = "Probability of what Trimester"
  )
A


# Model Prediction
B<-test.data1 %>%mutate(trim=predicted.classes)%>%
  mutate(prob = ifelse(trim == "3rd", 1, 0)) %>%
  ggplot(aes(CI, prob)) +geom_jitter(height = 0.01)+
  geom_point(alpha = 0.2) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  labs(
    title = "Logistic Regression Model \n Testing data set 40% \n Prediction", 
    x = "Cephalic Index",
    y = "Probability of what Trimester"
  )
B
```

## Plot showing both seen and unseen data

```{r}
library(ggpubr)
ggarrange(A,B,ncol = 2,nrow = 1)
```

