---
title: 'Determination of Normal CI in 2nd and 3rd Trimester: Sample population'
author: "Usman"
date: "2023-10-22"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    theme: journal
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

## **R Markdown**

# **Objective**
1. Estimation of fetal cephalic index among Hausa/Fulani pregnant women in Rano LGA, Kano State. 
2. Study the difference between 2nd and 3rd-trimester cephalic index values for both male and female
3. Assess possible association between sex and various obstetric parameters
4. Correlate the different fetal biometric parameters and gestational age to determine the best predictor

## **Inclusion criteria**:
- Singleton pregnancy. Dichorionic pregnancies with demise of 1 fetus in the first trimester were included
- Formal pregnancy ultrasound performed at WCH at 28 to 33 weeks, including standard fetal biometry. 
- Morphology ultrasound performed at 17 to 22 weeks’ gestation at WCH or another accredited images for review.


```{r}
library(readxl)
cephalic <- read_excel("~/Bluetooth Exchange Folder/sagagi2.xlsx",sheet = "Sheet2")
cephalic
```

## Figures, Code chunks, and Logistic regression tables

### **Figure 1**


```{r pressure}
library(tidyverse)
library(ggpubr)
#library(ggpubr) and library(tidyverse)
# Vertical and Horizontal median representation of both variables
#scatterplot showing correlation BPD and OFD at 2nd Trimester
meanbpd2<-cephalic%>%mutate(EG.A=str_replace_all(EG.A,"[+]","."),TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD",
  "2NDF"="2ND"))%>%filter(TRIMESTER=="2ND")%>%summarise(median_val=median(BPD))
meanbpd2

meanofd2<-cephalic%>%mutate(EG.A=str_replace_all(EG.A,"[+]","."),TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD",
"2NDF"="2ND"))%>%filter(TRIMESTER=="2ND")%>%summarise(median_val=median(OFD))
meanofd2

cephalic%>%mutate(EG.A=str_replace_all(EG.A,"[+]","."),TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  filter(TRIMESTER=="2ND")%>%ggplot(aes(BPD,OFD))+geom_point()+theme_minimal()+
  geom_hline(data= meanofd2, aes(yintercept = median_val),linetype="dashed")+
  geom_vline(data= meanbpd2, aes(xintercept = median_val),linetype="longdash")+geom_smooth(method = lm,se=FALSE,fullrange=TRUE)+
  scale_x_log10()+stat_cor(method = "spearman")
```

**Figure 1**: Scatter plot of OFD versus BPD at the morphology scan; points
have been ‘jittered’ horizontally and vertically to separate identical values; the dash lines are the median values for OFD and BPD; the solid blue line represents the correlation line

### **Figure 2**
```{r}
meanbpd3<-cephalic%>%mutate(EG.A=str_replace_all(EG.A,"[+]","."),
  TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  filter(TRIMESTER=="3RD")%>%
  summarise(median_val=median(BPD))

meanofd3<-cephalic%>%mutate(EG.A=str_replace_all(EG.A,"[+]","."),
                        TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD",
                        "2NDF"="2ND"))%>%filter(TRIMESTER=="3RD")%>%
                         summarise(median_val=mean(OFD))


cephalic%>%mutate(EG.A=str_replace_all(EG.A,"[+]","."),TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  filter(TRIMESTER=="3RD")%>%ggplot(aes(BPD,OFD))+
  geom_point()+theme_minimal()+
  geom_hline(data= meanofd3, aes(yintercept = median_val),linetype="dashed")+
  geom_vline(data= meanbpd3, aes(xintercept = median_val),linetype="longdash")+
  geom_smooth(method = lm,se=FALSE,fullrange=TRUE)+scale_x_log10()+scale_y_log10()+
  stat_cor(method = "spearman")

```

**Figure 2**: Scatter plot of OFD versus BPD at the growth scan; points
have been ‘jittered’ horizontally and vertically to separate identical values; the dash lines are the median values for OFD and BPD; the solid blue line represents the correlation line

### **Figure 3**

```{r}
#Scatter plot of the cephalic index at the growth scan versus the cephalic index at the morphology scan
ci_morho<-cephalic%>%mutate(EG.A1=str_replace_all(EG.A,"[+]","."),
  TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  mutate(CI3=(BPD/OFD)*100)%>%filter(TRIMESTER=="3RD")%>%
  select(TRIMESTER,CI3,EG.A1)
ci_growth<-cephalic%>%mutate(EG.A=str_replace_all(EG.A,"[+]","."),
  TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  mutate(CI2=(BPD/OFD)*100)%>%filter(TRIMESTER=="2ND")

ci_medianM<-tibble(ci_growth,CI3=ci_morho[-c(199:433),]$CI3)%>%
  summarise(median_val=median(CI2))

ci_medianG<-tibble(ci_growth,CI3=ci_morho[-c(199:433),]$CI3)%>%
  summarise(median_val=median(CI3))

tibble(ci_growth,CI3=ci_morho[-c(199:433),]$CI3)%>%ggplot(aes(CI2,CI3))+
  geom_point()+theme_minimal()+scale_x_log10()+scale_y_log10()+
  labs(x="CI at morphology scan",y="CI at growth scan")+
  geom_hline(data= ci_medianG, aes(yintercept = median_val),linetype="dashed")+
  geom_vline(data= ci_medianM, aes(xintercept = median_val),linetype="longdash")+
  geom_smooth(method = lm,se=FALSE,fullrange=TRUE)+stat_cor(method = "spearman")

```

**Figure 3**: Scatter plot of the cephalic index at the growth scan versus the cephalic index at the morphology scan, showing little difference between the 2 groups.


### **Figure 4**

```{r}
ci_growth1<-cephalic%>%mutate(EG.A=str_replace_all(EG.A,"[+]","."),
  TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  mutate(CI2=(BPD/OFD)*100)%>%filter(TRIMESTER=="2ND")%>%
  mutate(CI3=CI2)%>%select(TRIMESTER,CI3,EG.A)

ci_morpho<-cephalic%>%mutate(EG.A1=str_replace_all(EG.A,"[+]","."),
  TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  mutate(CI2=(BPD/OFD)*100)%>%filter(TRIMESTER=="3RD")%>%
  select(TRIMESTER,CI2,EG.A1)

ci_morpho$EG.A1<-as.numeric(ci_morpho$EG.A1)
ci_growth1$EG.A<-as.numeric(ci_growth1$EG.A)

new_data<-tibble(ci_growth1[,-1],ci_morpho[-c(199:433),])%>%
  mutate(gest_dif_wk= ci_morpho[-c(199:433),]$EG.A1-ci_growth1[,-1]$EG.A,ci_change=CI3-CI2)

new_data%>%ggplot(aes(gest_dif_wk,ci_change))+
  geom_point()+geom_smooth(method = lm,se=FALSE,fullrange=TRUE)+
  stat_cor(method = "spearman")+scale_y_continuous(limits = c(-10,10.3))+
  scale_x_continuous(limits = c(0,20))+geom_hline(yintercept = 0, linetype="dashed")+
  theme_minimal()+labs(x="difference in Gestation (week)", y= "Change in CI")

```

**Figure 4**: Scatter plot of the change in cephalic index between growth and morphology scans versus the corresponding change in
gestation age; the blue solid line shows the linear regression line; the dash tick line is a reference line for zero change in CI.

### **Figure 5**

```{r}
#comparing means and adding to our figure
my_comparisons<-list(c("2ND","3RD"))

cephalic%>%mutate(TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  ggplot(aes(TRIMESTER,CI))+geom_boxplot()+geom_point(position=position_jitter())+
  theme_minimal()+stat_compare_means(comparisons = my_comparisons)+facet_wrap(~SEX)

```

**Figure 5**: Comparison between group relative to gender

### **Figure 6**: using 'ggpairs()' 

```{r}
library(GGally)
#pairwise comparisons between morphology and growth variables
cephalic%>%mutate(TRIMESTER=recode(TRIMESTER,"2nd"="2ND","3rd"="3RD","2NDF"="2ND"))%>%
  ggpairs(columns=3:7,aes(col=TRIMESTER))+theme(panel.grid.major = element_blank(),
  axis.line = element_line(colour = "black"))+
  theme(strip.background = element_rect(fill = "white"),strip.placement = "outside")+
  theme(axis.text.x = element_blank())+scale_color_manual(values=c("blue","black"))+
  scale_fill_manual(values=c("blue","black"))
```

**Figure 6** Correlation matrix

### **Logistic regression analysis**

```{r}
#logistic regression analysis
cephalic$sex<-cephalic$SEX
cephalic$logit<-cephalic$TRIMESTER
cephalic[cephalic$logit=="2ND"|cephalic$logit=="2nd"|cephalic$logit=="2NDF",]$logit<-"0"
cephalic[cephalic$logit=="3RD"|cephalic$logit=="3rd",]$logit<-"1"
cephalic[cephalic$sex=="F",]$sex<-"0"
cephalic[cephalic$sex=="M",]$sex<-"1"
cephalic$logit<-as.factor(cephalic$logit)
glm(logit~sex, data=cephalic,family = "binomial")

summary(glm(logit~sex, data=cephalic,family = "binomial"))
summary(glm(logit~SEX+BPD+OFD+FL+AC, data=cephalic,family = "binomial"))
```


